Comparing Apples and Oranges in Youth Depression Treatments?

### 2024-09-22 revision1_reply_to_reviewers.qmd

---
title: "Comparing Apples and Oranges in Youth Depression Treatments? A Quantitative Critique of the Evidence Base and Guidelines"
subtitle: "Major Revision: Letter to Reviewers"
format: 
  docx:
    fig-width: 9
    fig-height: 7
    fig_caption: true
    tbl_caption: true
    number-sections: true
fig-cap-location: top
---

```{r, include=FALSE}
library(tidyverse)
library(truncnorm)
library(metafor)
library(meta)
library(kableExtra)
library(flextable)
library(magrittr)
library(officer)
library(RefManageR)
library(ftExtra)
library(dplyr)
library(broom)
library(effsize)
library(here)

#set source
load("my_environment.RData") #image that contains objects after main qmd has been run
```

@@

#### Comment 1.7

I could not see the rational for excluding the all female trials in the sex analyses. \[Comment 1.7\]

#### Response to 1.7

This is to not unduly bias the results. We now present the analyses with all studies included in the Supplemental Materials (see @tbl-all-studies-percfemale-results).

```{r, warning=FALSE, message = FALSE, echo=FALSE, ft.align="left"}
#| label: tbl-all-studies-percfemale-results
#| tbl-cap: "Percentage female at baseline across medication and psychotherapy studies: Results for sample including all female studies"
#| tbl-cap-location: top

@@

$$
Eff_A = Eff_B
$$

Since the AR is the same in each case, and one might conclude that we should be indifferent between the two treatments. However, it should be obvious that any rational agent, i.e. an agent who want so maximise benefit for themselves, would choose treatment A over B, if this were the data that they had to base their decision upon; indeed, nobody woudl choose a treatment with 40% probability recovery over one with 90% probability recovery? Unless of course there were a fundamental difference in the composition or circumstances of the trials.

All these considerations are very important and relevant to our paper and in the spirit of being very clear with the primary data, we have added, in addition to the  figure with the summary SMDs, this figure which presents the SMD of each study separately.

```{r, warning=FALSE, message = FALSE, echo=FALSE, ft.align="left"} #  #}

list_df_simulated[[1]] %>%  
  filter(arm_effect_size == "cohens_d_control") %>%  
  ggplot(aes(x = cohens_d, y = 0, color = as.factor(psy_or_med), size = baseline_n)) + 
  geom_jitter(height = 0.2, alpha = 0.7) + 
  geom_point(alpha = 0.7) + 
  labs(
    title = "Pre-Post Standardised Mean Differences (SMD) of Control Arms",
    x = "SMD",
    y = "",
    color = "Type of Study",
    size = "Baseline n") +
    scale_color_discrete(name = "Type of Study", labels = c("Medication", "Psychological Therapy")) +
      theme_minimal() +
      theme(
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank())
```

@@

#### Response to 3.10

We thank the reviewer for pointing this out. The key question is whether the baseline SDs are systematically different between medication and psychological therapy trials, i.e. whether this introduces a bias. If the differences are random, then this only contributes to noise. We are doing two things to address this.

First, we plot the mean differences (endpoint -- baseline) of the min-max transformed scores. As can be seen these are substantial, indicating that the difference is unlikely to stem simply from a difference in standard deviation calculation.

```{r, warning=FALSE, message = FALSE, echo=FALSE}
mean_diff_min_max_scores <- df_long_for_metan %>% 
  filter(arm_effect_size == "cohens_d_control") %>% 
  mutate(mean_dif = perc_post_mean - perc_baseline_mean) %>% 
  ggplot(aes(y = mean_dif, x = as.factor(psy_or_med), size = baseline_n)) + 
  geom_jitter(width = 0.2) + 
  ggtitle("Post-Pre Differences in Mean Min-Max Scores") + 
  ylab("Post-Pre Difference") +
  xlab("Type of Study") +
  scale_x_discrete(labels = c("0" = "Medication", "1" = "Psychological Therapy")) +   theme_minimal()

print(mean_diff_min_max_scores) 
png("mean_diff_min_max_scores.png")
mean_diff_min_max_scores 
dev.off()

@@

```

Second, we plot below the baseline variances (i.e. squared sds), which largely overlap, making it unlikely that differences in baseline SD is the cause of the difference in the outcome.

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# create baseline variances
df_long_for_metan$variances <- (df_long_for_metan$perc_baseline_sd)^2

plot_baseline_var <- df_long_for_metan %>% 
  filter(arm_effect_size == "cohens_d_control") %>% 
  mutate(mean_dif = perc_post_mean - perc_baseline_mean) %>% 
  ggplot(aes(y = variances, x = as.factor(psy_or_med), size = baseline_n)) + geom_jitter(width = 0.2) +  
  ggtitle("Post-Pre Differences in Study Baseline Variances") + 
  ylab("Baseline Variance") + 
  xlab("Type of Study") + 
  scale_x_discrete(labels = c("0" = "Medication", "1" = "Psychological Therapy")) + theme_minimal()

print(plot_baseline_var)
png("plot_baseline_var.png")
plot_baseline_var 
dev.off()

@@

```

Third, we conducted an addition analysis, restricting to studies with variances below 0.02 (see graph above).

```{r, warning=F}
#| label: fig-plot-means-var2
#| fig-cap: "Meta-analytic estimates of within-group changes: studies with variance <0.02" 
#| fig-cap-location: top

coef_and_se_means_var2_study <- coef_and_se_means_var2_study %>% 
  mutate(
    condition = case_when(condition == "medication_control" ~ "Medication Control", condition == "medication_active" ~ "Medication Active", condition == "psychotherapy_control" ~ "Psychotherapy Control",
condition == "psychotherapy_active" ~ "Psychotherapy Active", TRUE ~ as.character(condition)))

subtitle_text_hamd <- "Metanalytic estimates of within-group changes: HAM-D studies only"
plot_means_multi_var2 <- plot_means_function(coef_and_se_means_var2_study)
print(plot_means_multi_var2)
png("plot_means_multi_var2.png")
plot_means_multi_var2 # # dev.off() #  #

@@

2.  I would predict that few studies, if any, reported both the baseline and endpoint SD and the authors have apparently used various imputations based on so many assumptions to impute the pre-post SMD. \[Comment 3.11\]

#### Response to 3.11

Actually, we have endpoint SDs in 61 (69.3% of) trials. To address the reviewer's point, we have now reanalysed our data restricted to just those studies and as can be seen below, the same inferences hold as before. We include this figure now as Fig XXX in the Supplement.

```{r, warning=FALSE, message = FALSE, echo=FALSE}
#| label: fig-plot-means-var2
#| fig-cap: "Meta-analytic estimates of within-group changes: studies with post SD"
#| fig-cap-location: top

### 2024-09-22 multilevel_metaregression/manuscript_with_multilevel_metan.qmd
added cohen's d calculations

@@

As can be seen in tbl-baseline_results, mean age was `r round(results_met_age$TE_random[results_met_age$subgroup == 1][[2]],2)` (*SE* = `r``round(results_met_age$seTE_random[results_met_age$subgroup == 1][[2]],2)`) across psychotherapy trials and `r round(results_met_age$TE_random[results_met_age$subgroup == 0][[2]],2)` (*SE* = `r round(results_met_age$seTE_random[results_met_age$subgroup == 0][[2]],2)`) across medication trials, with no significant between group differences. There were no significant differences in mean age between modalities on further sensitivity analyses.

@@

# instances_count <- df_descr_psy_conditions_no_wl %>%
#   filter(!is.na(total_hours)) %>%
#   group_by(new_study_id, total_hours) %>%
#   summarise(count = n()) %>%
#   filter(count > 1) %>%
#   summarise(total_instances = n())

>>>>

```{r baseline-severity, echo=FALSE, message=FALSE, warning=FALSE}

mean_psych_overall <- 0.37
mean_med_overall <- 0.42
se_psych_overall <- 0.02
se_med_overall <- 0.01
n_psych_overall <- 49
n_med_overall <- 31

# Calculate standard deviation
sd_psych_overall <- se_psych_overall * sqrt(n_psych_overall)
sd_med_overall <- se_med_overall * sqrt(n_med_overall)

#formula with pooled sd
pooled_sd_overall <- sqrt(((sd_psych_overall^2 + sd_med_overall^2)/2))

# Calculate Cohen's d
cohens_d_overall <- (mean_psych_overall - mean_med_overall) / pooled_sd_overall

## without subclinical

mean_psych_sub <- 0.39
mean_med_sub <- 0.42
se_psych_sub <- 0.02
se_med_sub <- 0.01
n_psych_sub <- 41
n_med_sub <- 31

# Calculate standard deviation
sd_psych_sub <- se_psych_sub * sqrt(n_psych_sub)
sd_med_sub <- se_med_sub * sqrt(n_med_sub)

#formula with pooled sd
pooled_sd_sub <- sqrt(((sd_psych_sub^2 + sd_med_sub^2)/2))

# Calculate Cohen's d
cohens_d_sub <- (mean_psych_sub - mean_med_sub) / pooled_sd_sub
cohens_d_sub
cohens_d_overall
```

@@

As can be seen in tbl-baseline_results, mean age was `r round(results_met_age$TE_random[results_met_age$subgroup == 1][[2]],2)` (*SE* = ``` r``round(results_met_age$seTE_random[results_met_age$subgroup == 1][[2]],2) ```) across psychotherapy trials and `r round(results_met_age$TE_random[results_met_age$subgroup == 0][[2]],2)` (*SE* = `r round(results_met_age$seTE_random[results_met_age$subgroup == 0][[2]],2)`) across medication trials, with no significant between group differences. There were no significant differences in mean age between modalities on further sensitivity analyses.

@@

instances_count <- df_descr_psy_conditions_no_wl %>%
  filter(!is.na(total_hours)) %>%
  group_by(new_study_id, total_hours) %>%
  summarise(count = n()) %>%
  filter(count > 1) %>%
  summarise(total_instances = n())
